\input{configuration}

%% For wordcount run $texcount thesis.tex -inc -total

\begin{document}

\input{frontmatter}
\rmfamily
\tableofcontents{}
\listoffigures
\listoftables
\printnomenclature{}
\doublespacing

\chapter*{Abstract}
This thesis presents research into applications of interfacial double layers to electronic systems.
This research has two parts; layers on insulating surfaces and layers on conductors.

A double layer on an insulating surface offers the ability to convert energy between electric and fluid states.
Using this, I build streaming cell energy harvesters and determine their output characteristics.
I comment on the suitability of streaming cell technology as a means of powering smart water meters.

A double layer on a conductor causes interesting, but often unwanted, electrical behaviour in medical electrodes.
I use a newly developed model that captures this behaviour to characterise various fluids.
I characterise several fluids using this method.
I directly compare a standard solution of saline to the spinal fluid of a living animal.
Such a comparison is novel and important to medical implant designers.
I develop a way to create a fluid solution that better mimics the electrical properties of a given fluid.

\phantomsection\addcontentsline{toc}{chapter}{Abstract}

\chapter*{Acknowledgement}
\phantomsection\addcontentsline{toc}{chapter}{Acknowledgement}
Thanks go to Jonathan Scott (chief supervisor), Steve Newcombe (Waikato University's award winning glass blower) and Peter Single (senior engineer at Saluda Medical) for their time and resources.
Thank you to everyone who contributes to any of the following open source projects, from which I have benefitted enourmously,
\begin{itemize}
\item The Linux kernel
\item GNU software
\item The Gnome desktop environment
\item Inkscape
\item Gimp
\item Arch Linux
\item Tex and its derivative \LaTeX
\item Python
\item ngSpice
\end{itemize}

%% Part 1 - Introduction

\chapter{Introduction}
    \input{content/introduction/introduction.tex}

\part{Double Layers On Insulators}

    \label{part:doubleLayersOnInsulators}
        
    Part \ref{part:doubleLayersOnInsulators} deals with double layers formed on electrical insulators.
    Two examples of where this happens are the walls of plastic or glass containers.
    
    A practical application of double layers on insulators is to the field of energy harvesting.
    It was discussed in the introduction that double layers naturally form at liquid-solid interfaces.
    It was said that the double layer itself is a collection of charged ions.
    Can we collect this charge and convert it into electrical energy?

    In this part I explain how to harvest energy from water and glass without moving parts.
    I look at a smart application of a water such a harvester, create a power budget, and discuss the design of a practical harvester.

    \chapter{Harvesting Energy}
        \label{chap:harvestingEnergy}

        This chapter demonstrates an energy harvesting device capable of harvesting energy from flowing water with no moving parts.
        A `smart' use for such a harvester is discussed in chapter \ref{chap:wirelessWaterMetering}.
        I find the minimum power a harvester needs to harvest in order to be useful in chapter \ref{chap:energyRequirements}.
        And finally, in chapter \ref{chap:harvesterDesign}, I discuss the design of a harvester and comment on its feasibility.

        \input{content/pt1/01-PowerHarvesting/chapter}

    \chapter{Wireless Water Metering}
        \label{chap:wirelessWaterMetering}

        A method of power harvesting was put forward in chapter \ref{chap:harvestingEnergy} that used water flow as a source of energy.
        We expect the lack moving parts in that design will reduce or eliminate maintenance costs.
        Such a low maintenance harvester is ideal for water utility companies to measure and record water usage.
        This chapter discusses how such a water meter might be implemented.
        The content in this chapter is used as the basis for investigating the energy requirements of such a harvester.

        \input{content/pt1/02-WirelessWaterMeter/chapter}

    \chapter{Energy Requirements}
        \label{chap:energyRequirements}
        
        \section{Microcontroller}
            \input{content/pt1/02-Microcontrollers/chapter}
        
        \section{Data Transmission}
            \input{content/pt1/03-DataTransmission/chapter}

    \chapter{Harvester Design}
        \label{chap:harvesterDesign}
        \input{content/pt1/04-HarvesterDesign/chapter}

%% Part 3 - Double layers on conductors
\part{\label{part:doubleLayersOnConductors}Double layers on conductors}
    \chapter{Electrically Modelling an Interface}
        \input{content/pt2/01-ElectricalModelling/chapter}
    \chapter{Determining Interface Parameters}
        \input{content/pt2/02-InterfaceParameters/chapter}

%% Part 4 - Appendices
\part{Appendices}
    \appendix
    \chapter{Alternative Energy Harvester}
        \input{content/appendices/chargedWaterDrops/appendix}
    \chapter{Microprocessor Power Measurements}
        \input{content/appendices/microprocessorPowerMeasurements/appendix}
    \chapter{Streaming Cell Mesaurements}
        \input{content/appendices/streamingCellMeasurements/appendix}

\chapter*{Notes}
The text width of this thesis is \the\textwidth.


\bibliographystyle{plain}
\phantomsection\addcontentsline{toc}{chapter}{\bibname}
\bibliography{library}

\end{document}

