%!TEX root = ../../thesis.tex

Part \ref{part:doubleLayersOnInsulators} deals with double layers formed on electrical insulators.
Two examples of where this happens are the walls of plastic or glass containers.

A practical application of double layers on insulators is to the field of energy harvesting.
It was discussed in the introduction that double layers naturally form at liquid-solid interfaces.
It was said that the double layer itself is a collection of charged ions.
Can we collect this charge and convert it into electrical energy?

In this part I explain how to harvest energy from water and glass without moving parts.
I look at a smart application of a water such a harvester, create a power budget, and discuss the design of a practical harvester.

\chapter{Introduction}
  \label{chap:harvesterIntroduction}
  \section{Background}
  \section{Literature Review}

\chapter{The Streaming Cell Energy Harvester}
  \label{chap:harvestingEnergy}

  This chapter demonstrates an energy harvesting device capable of harvesting energy from flowing water with no moving parts.
  A `smart' use for such a harvester is discussed in chapter \ref{chap:wirelessWaterMetering}.
  I find the minimum power a harvester needs to harvest in order to be useful in chapter \ref{chap:energyRequirements}.
  And finally, in chapter \ref{chap:harvesterDesign}, I discuss the design of a harvester and comment on its feasibility.

  \input{content/pt1/01-PowerHarvesting/chapter}

\chapter{Applicability to Water Metering}
  \label{chap:wirelessWaterMetering}
  \section{Current Trends in Water Metering}
  \section{Appropriatenes of Streaming Cell Harvesters}

  A method of power harvesting was put forward in chapter \ref{chap:harvestingEnergy} that used water flow as a source of energy.
  We expect the lack moving parts in that design will reduce or eliminate maintenance costs.
  Such a low maintenance harvester is ideal for water utility companies to measure and record water usage.
  This chapter discusses how such a water meter might be implemented.
  The content in this chapter is used as the basis for investigating the energy requirements of such a harvester.
  \input{content/pt1/02-WirelessWaterMeter/chapter}

\chapter{Energy Requirements}
  \label{chap:energyRequirements}
  
  \section{Microcontrollers}
      \input{content/pt1/02-Microcontrollers/chapter}
  
  \section{Data Transmission}
      \input{content/pt1/03-DataTransmission/chapter}

\chapter{Discussion}

\chapter{Conclusion}